Efficient pedigree recording for fast population genetics simulation

### 2017-12-28 forwards_paper.tex
methods place holder
comment out bit about detailed timings
make clear what we are talking about re: performance

We ran all benchmarks on an Ubuntu Linux (version 16.04) system with two 2.6 Gigahertz Intel E5-2650 CPU with
hyperthreading enabled.
We ran one simulation at a time and the machine was under minimal load otherwise.
We used GNU parallel \citep{Tange2011a} to kill any simulation that did not finish within 72 hours,
and the Linux \texttt{time} command to record run time and peak memory usage of each replicate.
Our simulation scripts separately recorded the time spent in various steps of each run.

\krt{currently, we are not doing anything with the detailed timings.  I think that is fine, and that we can delete last
sentence of preceding paragraph.}

The total run times for simulations with selection are shown in Figure~\ref{fig:runtimes_selection}.  When tracking
neutral mutations (instead of the pedigree), run times increase dramatically with increasing region size ($4Nr = 4Nu$ for
these simulations).  With \fwdpy{}, simulations with $N=5 \times 10^4$ timed out for region sizes larger than $10^3$.
With \simupop{}, we were only able to run simulations with $N=10^3$ and region sizes up to $10^4$.  Pedigree tracking
dramatically improves run times for both \fwdpy{} and \simupop{} (right column of Figure~\ref{fig:runtimes_selection}).  For parameters that ran to completion when tracking neutral mutations and when tracking
the pedigree, the relative speedup due to pedigree tracking is up to $\approx 50$ fold
(Figure~\ref{fig:relative_speedup_selection}).  For \fwdpy{}, pedigree tracking reduces runtimes more for larger
population sizes and we observe increasing relative speedups as $4Nr$ increases for a given $N$
(Figure~\ref{fig:relative_speedup_selection}).  We saw a similar trend with $4Nr$ for \simupop{}, but the effect of $N$
could not be explored because the simulations tended to run out of available memory.
  Note the log scale of the x axis in Figure~\ref{fig:runtimes_selection}, which
partially obscures an important observation that the run times appear approximately linear in region size when pedigree
tracking.  The reason for this behavior is that the simulations are generating a constant number of new edges on
average
each generation as the number of new recombination breakpoints is Poisson distributed with expectation $\rho$.

@@

>>>>

We ran all benchmarks on an Ubuntu Linux (version 16.04) system with two 2.6 Gigahertz Intel E5-2650 CPU with
hyperthreading enabled.
We ran one simulation at a time and the machine was under minimal load otherwise.
We used GNU parallel \citep{Tange2011a} to kill any simulation that did not finish within 72 hours,
and the Linux \texttt{time} command to record run time and peak memory usage of each replicate.
%% Our simulation scripts separately recorded the time spent in various steps of each run.

%% \krt{currently, we are not doing anything with the detailed timings.  I think that is fine, and that we can delete last
%%sentence of preceding paragraph.}

The total run times for simulations with selection are shown in Figure~\ref{fig:runtimes_selection}.  When tracking
neutral mutations (instead of the pedigree), run times increase dramatically with increasing region size ($4Nr = 4Nu$ for
these simulations).  With \fwdpy{}, simulations with $N=5 \times 10^4$ timed out for region sizes larger than $10^3$.
With \simupop{}, we were only able to run simulations with $N=10^3$ and region sizes up to $10^4$.  Pedigree tracking
dramatically improves run times for both \fwdpy{} and \simupop{} (right column of Figure~\ref{fig:runtimes_selection}).  For parameters that ran to completion when tracking neutral mutations and when tracking
the pedigree, the relative speedup due to pedigree tracking is up to $\approx 50$ fold
(Figure~\ref{fig:relative_speedup_selection}).  For \fwdpy{}, pedigree tracking reduces runtimes more for larger
population sizes and we observe increasing relative speedups as $4Nr$ increases for a given $N$
(Figure~\ref{fig:relative_speedup_selection}).  We saw a similar trend with $4Nr$ for \simupop{}, but the effect of $N$
could not be explored because the simulations tended to run out of available memory.
  Note the log scale of the x axis in Figure~\ref{fig:runtimes_selection}, which
partially obscures an important observation that the run times appear approximately linear in region size when tracking
pedigree.  The reason for this behavior is that, when tracking pedigrees, the simulations are generating a constant number of new edges on
average
each generation as the number of new recombination breakpoints is Poisson distributed with expectation $\rho$.

@@

\krt{Are we doing a methods section?}

\subsection{Reagents list for the C++ implementation}

\begin{enumerate}
    \item \fwdpy{} 0.13.a0
    \item \fwdpp{} 0.5.7
    \item GNU Scientific Library 1.16
    \item pybind11 2.2.1
    \item GCC 4.8.5
\end{enumerate}

All of the above installed via conda.
